Carp\'{e} Data: Supporting Serendipitous Data Integration in Personal Information Management

### 2012-09-17 chi2012.tex
some rsults

\subsubsection{Study and task design}
To test these hypotheses, we designed a within-subjects lab-based A-B study in which participants were asked to perform two predefined tasks in separate timed trials, once with the DP interface (``condition A'') and once with a baseline interface (``condition B'' which we describe below).  We used upon the within-subjects A-B format in order to make both quantiative measurements about an individual's performance with DP (relative to baseline), but also to make qualitative observations of the processes used by each participant as they solved each task.  Each task was paired with one of two datasets, depending on condition, so that we could evaluate whether the participant's performance of completing a task differed with different datasets.  Conditions and interface order were fully counterbalanced among participants to eliminate inadvertent ordering and task bias.

@@

\subsection{Participant Recruitement}
We recruited participants through adverts posted near the University campus and e-mails to student mailing lists.  We screened participants to be at least 18 years of age, but did not filter or preferentially select participants based on any other criteria.  We took the first 20 that signed up and that successfully specified time slots that were available.   As described below in results, we ended up getting a large number of international students, who had varying proficiency in English.  Since all such participants had passed the requisite proficiency tests to enrol at University, we did not think this would substantially impact results.

\subsection{Experimental Setup}

The participants were given a standard desktop PC running MacOS X on a 22" monitor with a standard keyboard and mouse.  During the studies the Excel workbook and DP were initially set to full screen by default.  Each participant was asked to complete two tasks, one A and one B condition.  Each participant was given 10 minutes to complete a task.  In order to select which tasks the participant was given, we used a latin square to permutate the tasks and conditions.  In order to train the user on how to use our interface, they viewed a five-minute DataPalette tutorial video introducing its basic functionality, how to use histograms and maps, and linking instances.

@@

We recruited 20 participants (10 female) from the University of Southampton's international student mailing list.  All of the participants spoke English, and for 12 of them, English was not their first language.  Predominately the participants were studying various levels of Computer Science, and 5 of which studied other subjects.   On average each study took 40 mins to complete the tasks and exit surveys.

The information in Table \ref{tbl:data} was captured during the studies, so that we could statistically analyse the features used by the participants.

\begin{table*}[htdp]
\small
\begin{center}
\begin{tabular}{|r|p{0.3cm}|p{0.3cm}|p{0.3cm}|p{0.3cm}|p{0.3cm}|p{0.3cm}|p{0.3cm}|p{0.3cm}|p{0.3cm}|p{0.3cm}|p{0.3cm}|p{0.3cm}|p{0.3cm}|p{0.3cm}|p{0.3cm}|p{0.3cm}|p{0.3cm}|p{0.3cm}|p{0.3cm}|p{0.3cm}|}
\hline
Participants & 1 &2&3&4&5&6&7&8&9&10&11&12&13&14&15&16&17&18&19&20\\
\hline\hline
gender&f&m&f&f&f&f&f&m&f&f&m&f&m&f&m&m&m&m&m&m\\
A condition first task &y&y&y&n&n&y&n&n&y&y&n&n&y&y&n&n&n&n&n&n\\
\hline\hline
A Condition: task &4&2&3&4&3&&2&1&4&1&4&1&3&2&3&2&1&4&3&2\\
paper used&n&n&n&n&n&n&y&n&y&y&n&y&n&n&y&n&n&n&n&n\\
charts used &1&2&2&0&2&4&5&0&0&0&2&3&2&0&0&0&3&3&4&1\\
maps used &1&1&2&2&2&1&1&0&1&0&1&1&2&1&0&1&0&0&2&1\\ 
same as &4&3&3&6&0&2&0&2&1&0&6&5&6&2&7&0&6&7&7&7\\
number data sources  &4&3&4&4&4&3&4&3&4&4&3&3&5&4&5&4&4&5&5&4\\
chosen restaurant/uni &4.1&2.4&3.5&4.3&3.3&&2.4&1.4&4.2&1.6, \newline 1.3, \newline 1.2& 4.2& 1.4, \newline 1.5, \newline 1.3& 3.2& 2.1, \newline 2.3, \newline 2.6& 3.5, \newline 3.3&2.6, \newline 2.3, \newline 2.5&1.6, \newline 1.5, \newline 1.3&4.1&3.3&2.4, \newline 2.5, \newline 2.6\\
\hline\hline
 B Condition:task &3&&4&3&4&2&1&2&2&3&2&3&1&4&1&4&2&2&1&3\\
 paper used &n&n&n&n&n&n&y&n&y&y&n&y&n&y&y&n&n&y&n&n\\
 spreadsheet/website&&&&&&&&&&&&&&&&&&&&\\
 number of spreadsheets used&5/5&4/4&2/5&1.5&2/5&5/5&4/4&2/4&3/4&&2/4&4/5&2/4&4/5&2/4&2/5&4/4&4/4&0&5/5\\
 number of websites used&1&1&4&3&3&0&2&3&2&&1&2&3&1&2&3&2&0&4&0\\
 chosen restaurant/uni&3.5&&4.2&3.3&4.3&2.1&1.4, \newline 1.2, \newline 1.6&2.1&2.3, \newline 2.6, \newline 2.5&3.5&2.2, \newline 2.3, \newline 2.6&3.5&1.1&4.1&1.6&4.2&2.6, \newline 2.3, \newline 2.5&2.3, \newline 2.5, \newline 2.6&1.1, \newline 1.3,  \newline 1.4&3.5\\
 \hline
\end{tabular}
\end{center}
\caption{Participants data.} \label{tbl:data}
\normalsize
\end{table*}

In the A condition, where participants used DataPalette to complete their tasks they used it to collate the data on-screen, and compare multiple values directly. About half of the observed users actually used the instance combination feature of DataPalette, with the other half able to successfully make decisions without using it. Those who tried to use it were largely successful, and were able to therefore use the global brush-highlighting to simultaneously compare different statistics. Most users instantly opened up all data sources into individual boxes, in order to determine which sources contained specific statistics. The result of this technique is that the users instantly filled their screen, and ran out of space for plots and maps.

In the B conditions, participants were given the option of using websites and or a spreadsheet to make their choice.  Participants who opted to use websites to make their decisions repeatedly encountered a number of difficulties and issues. In particular, there was significant confusion over the meaning of data points, for example the percentage-based rankings of university teaching by the Times Higher Education. There were numerous usability issues in using websites, particularly in searching for university courses and course requirements. Similarly, participants found it difficult to understand the locations of restaurants in cities that they were not familiar, and could not find a way to plot the shortlist of restaurants onto the same map using the websites. Overall, the users that spent the most time using websites exhibited significant frustration attempting to locate and compare different criteria (such as course entry requirements, or restaurant locations), and all ended up falling back to using the supplied spreadsheets.

>>>>

\subsubsection{Study and task design}
To test these hypotheses, we designed a within-subjects lab-based A-B study in which participants were asked to perform two predefined tasks in separate timed trials, once with the DP interface (``condition A'') and once with a baseline interface (``condition B'' which we describe below).  We used upon the within-subjects A-B format in order to make both quantiative measurements about an individual's performance with DP (relative to baseline), but also to make qualitative observations of the processes used by each participant as they solved each task.  Each task was paired with one of two datasets, depending on condition, so that we could evaluate whether the participant's performance of completing a task differed with different datasets.  Conditions and interface order were fully counterbalanced among participants to eliminate ordering and task bias.

@@

\subsection{Participant Recruitment}
We recruited participants through adverts posted near the University campus and e-mails to student mailing lists.  We screened participants to be at least 18 years of age, but did not filter or preferentially select participants based on any other criteria.  We took the first 20 that signed up and that successfully specified time slots that were available.   As described below in results, we ended up getting a large number of international students, who had varying proficiency in English.  Since all such participants had passed the requisite proficiency tests to enrol at University, we did not think this would substantially impact results.

Participants were offered a small gift certificate from an online retailer in exchange for their time.

\subsection{Experimental Setup}
The participants were given a standard desktop PC running MacOS X on a 22" monitor with a standard keyboard and mouse.  During the studies the Excel workbook and DP were initially set to full screen by default.  Each participant was asked to complete two tasks, one A and one B condition.  Each participant was given a maximum of 10 minutes to complete a task, but were told that if they finished earlier they could tell us and move on to the next condition.  In order to select which tasks the participant was given, we used a latin square to permutate the tasks and conditions.  In order to train the user on how to use our interface, they viewed a five-minute DataPalette tutorial video introducing its basic functionality, how to use histograms and maps, and linking instances.

@@

Twenty-six (26) participants responded, from which we selected the first 20.  Ten participants were female.  All of the participants spoke English, but 12 of them were non-native English speakers.  Seven were non-students; out of the remaining 13, eight studied computer science and 5 studied other subjects.   On average, each study took 40 mins to complete.

The information in Table \ref{tbl:data} summarizes the metrics that were captured of the trials for each participant. We analyze this data next.

\begin{table*}[htdp]
\small
\begin{center}
\begin{tabular}{|r|p{0.3cm}|p{0.3cm}|p{0.3cm}|p{0.3cm}|p{0.3cm}|p{0.3cm}|p{0.3cm}|p{0.3cm}|p{0.3cm}|p{0.3cm}|p{0.3cm}|p{0.3cm}|p{0.3cm}|p{0.3cm}|p{0.3cm}|p{0.3cm}|p{0.3cm}|p{0.3cm}|p{0.3cm}|p{0.3cm}|}
\hline
Participants & 1 &2&3&4&5&6&7&8&9&10&11&12&13&14&15&16&17&18&19&20\\
\hline\hline
gender&f&m&f&f&f&f&f&m&f&f&m&f&m&f&m&m&m&m&m&m\\
A condition first task &y&y&y&n&n&y&n&n&y&y&n&n&y&y&n&n&n&n&n&n\\
\hline\hline
A Condition: task &4&2&3&4&3&&2&1&4&1&4&1&3&2&3&2&1&4&3&2\\
paper used&n&n&n&n&n&n&y&n&y&y&n&y&n&n&y&n&n&n&n&n\\
charts used &1&2&2&0&2&4&5&0&0&0&2&3&2&0&0&0&3&3&4&1\\
maps used &1&1&2&2&2&1&1&0&1&0&1&1&2&1&0&1&0&0&2&1\\ 
same as used &4&3&3&6&0&2&0&2&1&0&6&5&6&2&7&0&6&7&7&7\\
\# data sources used &4&3&4&4&4&3&4&3&4&4&3&3&5&4&5&4&4&5&5&4\\
chosen restaurant/uni &4.1&2.4&3.5&4.3&3.3&&2.4&1.4&4.2&1.6, \newline 1.3, \newline 1.2& 4.2& 1.4, \newline 1.5, \newline 1.3& 3.2& 2.1, \newline 2.3, \newline 2.6& 3.5, \newline 3.3&2.6, \newline 2.3, \newline 2.5&1.6, \newline 1.5, \newline 1.3&4.1&3.3&2.4, \newline 2.5, \newline 2.6\\
\# factors in choice &-&-&-&-&-&-&-&-&-&-&-&-&-&-&-&-&-&-&-&-\\
\hline\hline
 B Condition:task &3&&4&3&4&2&1&2&2&3&2&3&1&4&1&4&2&2&1&3\\
 paper used &n&n&n&n&n&n&y&n&y&y&n&y&n&y&y&n&n&y&n&n\\
 spreadsheet/website&&&&&&&&&&&&&&&&&&&&\\
 number of spreadsheets used&5/5&4/4&2/5&1.5&2/5&5/5&4/4&2/4&3/4&&2/4&4/5&2/4&4/5&2/4&2/5&4/4&4/4&0&5/5\\
 number of websites used&1&1&4&3&3&0&2&3&2&&1&2&3&1&2&3&2&0&4&0\\
 chosen restaurant/uni&3.5&&4.2&3.3&4.3&2.1&1.4, \newline 1.2, \newline 1.6&2.1&2.3, \newline 2.6, \newline 2.5&3.5&2.2, \newline 2.3, \newline 2.6&3.5&1.1&4.1&1.6&4.2&2.6, \newline 2.3, \newline 2.5&2.3, \newline 2.5, \newline 2.6&1.1, \newline 1.3,  \newline 1.4&3.5\\
 \# factors in choice &-&-&-&-&-&-&-&-&-&-&-&-&-&-&-&-&-&-&-&-\\
 \hline
\end{tabular}
\end{center}
\caption{Participants data.} \label{tbl:data}
\normalsize
\end{table*}

\subsection{Task performance}

We used three metris to measure of task performance: time to completion, the number of factors used in the choice, and the number of candidates

\subsubsection{Efficiency: Time taken}
Participants spent an average of 8.9 minutes performing the tasks overall, with DP (condition A) taking slightly less time ($M=8.7$min, $SD=1.79$) per trial than condition B (the web browser/Excel interface) ($M=9.0$min; $SD=1.44$).  An ANOVA test of time taken by interface condition blocking by participant did not find a interface use a significant predictor of task completion time. A comparison of times taken are visible in Figure \ref{fig:time}.

Although across trials, restaurant selection task took on average slightly less time ($M=8.8$min,$SD=1.58$) than the university trials ($M=8.9$min,$SD=1.66$), the aforementioned ANOVA test found task type to not be a significant predictor either.

\subsubsection{Thoroughness: Factors weighed in final choice}
A second metric was the number of factors the participant considered in making their final choice. To determine this, we asked participants to tell us precisely why their choice was made, and matched their justifications to properties in the data.  This analysis revealed a significant effect of interface on number of factors weighed ($F(1,19)=3.95, p<0.10$); as visible in Figure \ref{fig:factors}, participants in the DP condition listed an average of 5.2 choices in their decision ($SD=1.7$) versus 4.5 ($SD=1.73$) in the standard condition.

\subsubsection{Diversity: Data sources consulted}
A third metric we used to gauge how well the participant performed their tasks is the number of different data sources the participant consulted during the trial.

\subsubsection{Effort: External cognition}
Although we did not have an explicit measure of effort for this experiment (such as NASA's TLX \cite{tlx}) the use of paper notes provided an indicator of how much participants had to keep track of during the study.  We found that participants took notes less on average (25\%) with the DP interface than the standard interface (35\%), a difference which approached significance in an ANOVA blocked by participant ($F(1,19)= 8.22$, $p\approx0.10$).

\subsection{Condition A (DP)}
In the A condition, where participants used DataPalette to complete their tasks they used it to collate the data on-screen, and compare multiple values directly. About half of the observed users actually used the instance combination feature of DataPalette, with the other half able to successfully make decisions without using it. Those who tried to use it were largely successful, and were able to therefore use the global brush-highlighting to simultaneously compare different statistics. Most users instantly opened up all data sources into individual boxes, in order to determine which sources contained specific statistics. The result of this technique is that the users instantly filled their screen, and ran out of space for plots and maps.

\subsection{Condition B (control}
In the B conditions, participants were given the option of using websites and or a spreadsheet to make their choice.  Participants who opted to use websites to make their decisions repeatedly encountered a number of difficulties and issues. In particular, there was significant confusion over the meaning of data points, for example the percentage-based rankings of university teaching by the Times Higher Education. There were numerous usability issues in using websites, particularly in searching for university courses and course requirements. Similarly, participants found it difficult to understand the locations of restaurants in cities that they were not familiar, and could not find a way to plot the shortlist of restaurants onto the same map using the websites. Overall, the users that spent the most time using websites exhibited significant frustration attempting to locate and compare different criteria (such as course entry requirements, or restaurant locations), and all ended up falling back to using the supplied spreadsheets.
